---
title: "Re-pull the Sentinel 2 band data using the volunteer points"
author: "ROSSyndicate"
date: "2024-03-04"
output: html_document
---

# Purpose

This script uses the volunteer point locations exported from
`1_data_download.Rmd` which have been manually uploaded as a feature collection
to the `ee-ross-superior` project in Earth Engine.

## R/Python Setup

Load (and install, if needed) the necessary packages for this script.

```{r load-r-pkgs}
# get all the functions from the src folder and load them
source_files <- list.files("src", full.names = T)
invisible(lapply(source_files, source))

libs <- c("reticulate", "tidyverse")

invisible(lapply(libs, package_loader))
```

Use the conda environment if the env folder is present, otherwise create the
environment.

```{r conda-env}
if (!dir.exists("env")) {
  source("pySetup.R")
} else {
  use_condaenv(file.path(getwd(), "env"))
}
```

Load python modules and authenticate/initialize EE.

```{python}
import ee
import imp
imp.load_source("re_pull", "eePlumB/2_data_ingestion/re_pull_functions.py")
import re_pull as rp

ee.Authenticate()
ee.Initialize(project = "ee-ross-superior")
```

## Sort labels

Here, we'll load the entire label list, then filter for only the Sen2 labels

```{python}
labels = ee.FeatureCollection("projects/ee-ross-superior/assets/labels/collated_label_data_v2023-07-20")

labels_sen = labels.filter(ee.Filter.eq("mission", "SEN2"))
```

Get the unique dates of images from this list and assign the date to the
`system:time_start` parameter.

```{python}
dates_sen = labels_sen.aggregate_array("date").distinct()

labels_sen_dt = labels_sen.map(rp.set_date)
```

## Get feature collection ready

Filter the Sentinel 2 stack, then export the scene metadata

```{python}
# filter stack for desired tiles
TILES = ee.List(['15TWN', '15TXN', '15TYN', '15TWM', '15TXM', '15TYM'])

# sentinel stack, filtered
s2 = (ee.ImageCollection('COPERNICUS/S2_SR_HARMONIZED')
  .filter(ee.Filter.inList('MGRS_TILE', TILES)))

export_sen_meta = (ee.batch.Export.table.toDrive(
  collection = s2,
  description = 'SEN2_image_metadata',
  folder = 'EE_output',
  fileNamePrefix = 'SEN2_image_metadata',
  fileFormat = 'csv'))
  
#export_sen_meta.start()

```

And then apply the scaling factors and QA bits to the stack

```{python}
s2 = (s2
  .map(rp.applyScaleFactors_S2)
  .map(rp.apply_sat_defect_mask)
  .map(rp.add_qa_info_s2)
  .map(rp.flag_cirrus_opaque)
  .map(rp.addImageDate_S2))

```

## Export location information by dates

```{python}
# loop and export additional data
for i in range(dates_sen.length().getInfo()):
  one_date = dates_sen.get(i)
  print(one_date.getInfo())
  one_dt = ee.Date(one_date)
  dt_label = labels_sen_dt.filterDate(one_dt, one_dt.advance(1, 'day'))
  one_image_bands = (s2
    .filterDate(one_dt, one_dt.advance(1, 'day'))
    .select(["SR_.*"])
    .mean())
  one_image_qa = (s2
    .filterDate(one_dt, one_dt.advance(1, 'day'))
    .select(['MSK_CLDPRB', 'MSK_SNWPRB', 'cirrus', 'opaque', 'dark_pixel', 'cloud_shadow', 'water',
    'low_prob_cloud', 'med_prob_cloud', 'hi_prob_cloud', 'cirrus_scl', 'snow_ice'])
    .max())
  
  #define bands to extract and reduce regions
  bandsOut = one_image_bands.addBands(one_image_qa)

  combinedReducer = (ee.Reducer.median().unweighted()
                      .forEachBand(bandsOut.select("SR_.*"))
    .combine(ee.Reducer.max().unweighted()
      .forEachBand(bandsOut.select(['MSK_CLDPRB', 'MSK_SNWPRB', 'cirrus', 'opaque', 'dark_pixel', 'cloud_shadow', 'water',
    'low_prob_cloud', 'med_prob_cloud', 'hi_prob_cloud', 'cirrus_scl', 'snow_ice'])), sharedInputs = False)
    )
    
  # Collect median reflectance and occurance values
  # Make a cloud score, and get the water pixel count
  data = (bandsOut
    .reduceRegions(
      collection = dt_label,
      reducer = combinedReducer,
      scale = 30,
      crs = one_image_bands.geometry().projection().crs()))
  
  image_date_export = (ee.batch.Export.table.toDrive(
    collection = data,
    description = 'SEN2_' + one_date.getInfo(),
    folder = 'eePlumB_additional_band_data',
    fileNamePrefix = 'SEN2_' + one_date.getInfo() + '_additional_vars_v2024-04-25',
    fileFormat = 'csv'))
    
  image_date_export.start()

```
