---
title: "Move labels to repo"
author: "ROSSyndicate"
date: "2024-03-07"
output: html_document
---

```{r setup, echo = F}
# get all the functions from the src folder and load them
source_files <- list.files("src", full.names = T)
invisible(lapply(source_files, source))

# list/load/download the packages needed for this script
packages <- c('tidyverse',
              'googledrive')
invisible(lapply(packages, package_loader))
```

# Purpose

This script moves the collated labels datasets from the Drive folder to the local repository.

Authenticate Drive - these files are located in the ROSS account. Reach out for credentials.

```{r}
drive_auth(email = "therossyndicate@gmail.com")
```

Also, identify the version date we're interested in:

```{r}
# date for original collation
og_collation_date = "2023-07-20"
# date for secondary collation
add_data_date = "2024-04-25"
```

## Download collated files from eePlumB users

This file was created in `eePlumB/2_data_ingestion/1_data_download.Rmd`.

Find the collated labels data file for our version date, download from drive if
it is not in the repository

```{r}
# make the user-exported data collation file path
collated_fp <- paste0("data/labels/collated_label_data_v",
                      og_collation_date,
                      ".csv")

if (file.exists(collated_fp)) {
  message("collated file with provided collation date is already stored locally")
} else {
  # grab all the file info from the collated folder
  collated = drive_ls(path = 'eePlumB_collated_data')
  # get the right version date
  collated_version = collated[grepl(og_collation_date, collated$name),]
  # grab just the labels (not the validation data)
  labels = collated_version[grepl('label', collated_version$name),]
  collated_file <- retrieve_drive_file(driveid = labels$id)
  #save the file locally
  write_csv(collated_file, collated_fp)
}

```

## Download the files created using 'data_re_pull' scripts in \`eePlumB/2_data_ingestion\`

Make a list of the files in the drive and download them to a tmp folder, if the collated
file of the additional data doesn't exist. In this case, we'll save each mission group
as its own file, because the colnames are different and this will save us some
munging later.


```{r}
# get a list of the added band data
add_data_list <- drive_ls(path = "eePlumB_additional_band_data", pattern = "_v2024-04-25.csv")

process_new_labels_by_mission <- function(mission_text) {
  txt_for_grepl <- str_replace(mission_text, "_", "|")
  mission_files <- add_data_list %>% 
    filter(grepl(txt_for_grepl, name))
  # collate labels per mission
  labels <- map(mission_files$id,
                retrieve_drive_file) %>% 
    list_rbind()
  add_data_collation_fp <- paste0("data/labels/collated_labels_with_additional_data_",
                                  mission_text,
                                  "_v",
                                  add_data_date,
                                  ".csv")
  
  # save the file locally
  write_csv(labels, add_data_collation_fp)
}

walk(c("LS5_LS7", "LS8_LS9", "SEN2"),
     process_new_labels_by_mission)
```

Also grab the scene metadata for the data

```{r}
# get a list of the metadata files from Drive
metadata_list <- drive_ls(path = "EE_output", pattern = "image_meta")

process_metadata_by_mission <- function(mission_text) {
  txt_for_grepl <- str_replace(mission_text, "_", "|")
  mission_files <- metadata_list %>% 
    filter(grepl(txt_for_grepl, name))
  # collate labels per mission
  labels <- map(mission_files$id,
                retrieve_drive_file) %>% 
    list_rbind()
  add_data_collation_fp <- paste0("data/labels/collated_scene_metadata_",
                                  mission_text,
                                  "_v",
                                  add_data_date,
                                  ".csv")
  
  # save the file locally
  write_csv(labels, add_data_collation_fp)
}

walk(c("LS5_LS7", "LS8_LS9", "SEN2"),
     process_metadata_by_mission)
```

