---
title: "Shoreline Contamination Workflow"
author: "ROSSyndicate"
date: "2023-07-18"
output: html_document
editor_options:
  markdown:
    wrap: 80
---

# Purpose

This script creates a quick visualization of user-labeled shoreline
contamination (SC) and creates a new shapefile that has been edited to remove
areas where SC has been detected so that we can add some uncertainty measures
later in the model.

```{r setup, echo = FALSE}
packages <- c('tidyverse',
              'sf',
              'mapview')

package_loader <- function(x) {
    if (x %in% installed.packages()) {
      library(x, character.only = TRUE)
    } else {
      install.packages(x)
      library(x, character.only = TRUE)
    }
}

lapply(packages, package_loader)
```

## Load the labels file into the environment

```{r}
labels_file <- list.files('data/labels/', full.names = T)
labels <- read_csv(labels_file)
```

## Load the aoi into the environment

```{r}
aoi <- st_read('data/aoi/Superior_AOI_modeling.shp')
```

## Grab the shoreline contamination labels and plot on ma Filter for 'shorelineContamination'

```{r}
sc_labels <- labels %>%
  filter(class == 'shorelineContamination')
```

The crs is WGS 84, so let's save these as a sf

```{r}
sc <- st_as_sf(sc_labels, coords = c('lon', 'lat'), crs = 'EPSG:4326')

#and then convert to the crs of the aoi
aoi_crs <- st_crs(aoi)

sc <- st_transform(sc, aoi_crs)
```

And plot to see where they are in conjunction with the shapefile

```{r}
ggplot() +
  geom_sf(data = aoi) +
  geom_sf(data = sc, color = 'red') +
  theme_minimal()
```

As expected, this has pretty good coverage, especially since there are `r
nrow(sc)` labels.

Let's buffer these points by 60m and snip the polygon to account for these
areas of contamination.

```{r}
sc_buff <- st_buffer(sc, dist = 60)
# then perform a union so that there's only one feature
sc_buff_1 <- st_union(sc_buff)
# check to make sure
mapview(aoi) +
  mapview(sc_buff_1)
```

And now remove those areas from the polygon

```{r}
# make geometries valid
aoi <- st_make_valid(aoi)
sc_buff_1 <- st_make_valid(sc_buff_1)

# remove areas of sc_buff from original aoi
aoi_no_sc <- st_difference(aoi, sc_buff_1)

mapview(aoi_no_sc)
```

And let's remove the tiny areas that are created that are not part of the large
polygon.

```{r}
# get indiv polygons
aoi_no_sc_all <- st_cast(aoi_no_sc, 'POLYGON')
# calculate the area for each
aoi_no_sc_all$area <- st_area(aoi_no_sc_all)
# sort them by size and only grab the largest one
aoi_no_sc_one <- aoi_no_sc_all %>%
  arrange(-area) %>%
  slice(1) %>%
  st_make_valid(.)

mapview(aoi_no_sc_one)
```

And now trim the SC layer to the modeling AOI.

```{r}
sc_buff_export <- sc_buff_1 %>%
  st_intersection(., aoi) %>%
  st_make_valid(.)

mapview(sc_buff_export)
```

And finally, write the shapefiles

```{r}
st_write(sc_buff_export, 'data/aoi/Superior_shoreline_contamination.shp', append = F)
st_write(aoi_no_sc, 'data/aoi/Superior_AOI_minus_shoreline_contamination.shp', append = F)
```

```{r, echo = F}
knitr::wrap_rmd('modeling/01_Shoreline_Contamination.Rmd', width = 80, backup = NULL)
```
