---
title: "03_Shoreline_Contamination.Rmd"
author: "ROSSyndicate"
date: "2023-07-18"
output: html_document
---

# Purpose

This script creates a quick visualization of user-labeled shoreline
contamination (SC) and creates a new shapefile that has been edited to remove
areas where SC has been detected so that we can add some uncertainty measures
later in the model.


```{r setup, echo = FALSE}
packages <- c('tidyverse',
              'googledrive',
              'sf',
              'mapview')

package_loader <- function(x) {
    if (x %in% installed.packages()) {
      library(x, character.only = TRUE)
    } else {
      install.packages(x)
      library(x, character.only = TRUE)
    }
}

lapply(packages, package_loader)

# point to directories
shp_dir = 'data/aoi/'
tmp_dir = 'tmp/'

#create temporary directory for drive downloads
try(dir.create('tmp'))

v_date = '2023-07-19'
```

# Download collated files from eePlumB users

Authenticate Drive

```{r}
drive_auth()
```

Some helper functions

```{r}
drive_downloader = function(driveid, drivename, dest) {
  drive_download(as_id(driveid), path = file.path(dest, drivename), overwrite = T)
}

driveUp = function(dumpid, directory, filename) {
  drive_upload(file.path(directory, filename), filename, overwrite = T, path = as_id(dumpid))
}
```

Find the collated labels data file for our version date

```{r}
collated = drive_ls(path = 'eePlumB_collated_data')
collated_version = collated[grepl(v_date, collated$name),]
labels = collated_version[grepl('label', collated_version$name),]
```

Download the file to the temporary folder and then read into environment

```{r}
drive_downloader(labels$id[1], labels$name[1], tmp_dir)
coll_labels <- read_csv(file.path(tmp_dir, labels$name[1]))
```

## Load the aoi

```{r}
aoi <-  st_read('data/aoi/Superior_AOI_modeling.shp')
```


## Grab the shoreline contamination labels and plot on map

Filter for 'shorelineContamination'

```{r}
sc_labels <- coll_labels %>%
  filter(class == 'shorelineContamination')
```

The crs is WGS 84, so let's save these as a sf

```{r}
sc <- st_as_sf(sc_labels, coords = c('lon', 'lat'), crs = 'EPSG:4326')

#and then convert to the crs of the aoi
aoi_crs <- st_crs(aoi)

sc <- st_transform(sc, aoi_crs)
```

And plot to see where they are in conjunction with the shapefile

```{r}
ggplot() +
  geom_sf(data = aoi) +
  geom_sf(data = sc, color = 'red') +
  theme_minimal()
```

As expected, this has pretty good coverage, especially since there are `r
nrow(sc)` labels.

Let's buffer these points by 60m and snip the polygon to account for these
areas of contamination.

```{r}
sc_buff <- st_buffer(sc, dist = 60)
# then perform a union so that there's only one feature
sc_buff_1 <- st_union(sc_buff)
# check to make sure
mapview(aoi) +
  mapview(sc_buff_1)
```

And now remove those areas from the polygon

```{r}
# make geometries valid
aoi <- st_make_valid(aoi)
sc_buff_1 <- st_make_valid(sc_buff_1)

# remove areas of sc_buff from original aoi
aoi_no_sc <- st_difference(aoi, sc_buff_1)

mapview(aoi_no_sc)
```

And finally, write the shapefile

```{r}
st_write(sc_buff_1, 'data/aoi/Superior_shoreline_contamination.shp', append = F)
st_write(aoi_no_sc, 'data/aoi/Superior_AOI_minus_shoreline_contamination.shp', append = F)
```


And clean up temp folder

```{r}
unlink('tmp', recursive = F)
```


```{r, echo = F}
knitr::wrap_rmd('modeling/01_Shoreline_Contamination.Rmd', width = 80, backup = NULL)
```
