---
title: "eePlumB Yugo Model"
author: "ROSSyndicate"
date: "2023-07-06"
output: html_document
---

# Set up

```{r setup, echo = F}
libs = c('reticulate', 'tidyverse')

package_loader <- function(x) {
    if (x %in% installed.packages()) {
      library(x, character.only = TRUE)
    } else {
      install.packages(x)
      library(x, character.only = TRUE)
    }
}

lapply(libs, package_loader)
```

## Activate conda environment

Check for virtual environment and activate, otherwise, set up virtual
environment.

```{r conda env}
if (!dir.exists("env")) {
  source("pySetup.R")
} else {
  use_condaenv(file.path(getwd(), "env"))
}
```

And then import the needed modules

```{python}
import ee
import time
```

## GEE Setup

In order to access the GEE API, we need to authenticate our identity. To do
this, you must type `earthengine authenticate` in the terminal window. When
your browser states 'You are now authenticated with the gcloud CLI', the
authentication is complete. This authentication is valid for 7 days.

Now, we need to initialize our GEE session.

```{python}
ee.Initialize(project = 'ee-ross-superior')
```

# Import assets

## Load label data

```{python}
classValues = (['cloud',
  'openWater',
  'lightNearShoreSediment',
  'offShoreSediment',
  'darkNearShoreSediment'])

labels = (ee.FeatureCollection("projects/ee-ross-superior/assets/labels/collated_label_data_v2023-05-05")
  .filter(ee.Filter.inList('class', classValues)))

```


## Load the image collection

```{python}
# list bandnames; in part for L7 bands to match l8/9
bn457 = ['SR_B1', 'SR_B2', 'SR_B3', 'SR_B4', 'SR_B5', 'SR_B7', 'QA_RADSAT']
bn89 = ['SR_B2', 'SR_B3', 'SR_B4', 'SR_B5', 'SR_B6', 'SR_B7', 'QA_RADSAT']

# load Landsat 4, 5, 7, 8, 9 Surface Reflectance and rename bands
l9 = ee.ImageCollection("LANDSAT/LC09/C02/T1_L2")
l8 = ee.ImageCollection("LANDSAT/LC08/C02/T1_L2")
l7 = ee.ImageCollection('LANDSAT/LE07/C02/T1_L2')
l5 = ee.ImageCollection('LANDSAT/LT05/C02/T1_L2')
l4 = ee.ImageCollection('LANDSAT/LT04/C02/T1_L2')

# harmonize names
l457 = l4.merge(l5).merge(l7).select(bn457, bn89)
l89 = l8.merge(l9).select(bn89)

# merge to one collection
ls = l457.merge(l89)

# filter stack for desired PRs
ROWS = ee.List([27, 28])

ls = (ls
  .filter(ee.Filter.eq('WRS_PATH', 26))
  .filter(ee.Filter.inList('WRS_ROW', ROWS)))
```


## Apply scaling factors

```{python}
def applyScaleFactors(image):
  opticalBands = image.select('SR_B.').multiply(0.0000275).add(-0.2)
  return image.addBands(opticalBands, None, True)

ls = ls.map(applyScaleFactors)

```

## Load AOIs and clip stack

```{python}
aoi1 = ee.FeatureCollection('projects/ee-ross-superior/assets/tiledAOI/SuperiorAOI_1')
aoi2 = ee.FeatureCollection('projects/ee-ross-superior/assets/tiledAOI/SuperiorAOI_2')
aoi3 = ee.FeatureCollection('projects/ee-ross-superior/assets/tiledAOI/SuperiorAOI_3')
aoi4 = ee.FeatureCollection('projects/ee-ross-superior/assets/tiledAOI/SuperiorAOI_4')
aoi5 = ee.FeatureCollection('projects/ee-ross-superior/assets/tiledAOI/SuperiorAOI_5')

all_aois = (ee.FeatureCollection(aoi1)
  .merge(aoi2)
  .merge(aoi3)
  .merge(aoi4)
  .merge(aoi5))

all_aois = all_aois.union()

# clip images to aoi
def clip(image):
  return image.clip(all_aois.geometry())

ls_aoi = ls.map(clip)

```


#  Create Gradient Boost Tree Model

## Define the input features and output labels

```{python}
inputFeatures = ["SR_B2", "SR_B3", "SR_B4", "SR_B5", "SR_B6", "SR_B7"]
outputLabel = "class"
```

## Remap the label values to a 0-based sequential series.

```{python}
remapValues = ee.List.sequence(0, 4)
labels_map = labels.remap(ee.List(classValues), remapValues, ee.String(outputLabel))

def class_to_byte(feature):
  byteValue = ee.Number(feature.get(outputLabel)).toByte()
  return feature.set('byte_property', byteValue)

labels_map = labels_map.map(class_to_byte)

print('Filtered labels:', labels_map.getInfo())
```

## Split labels into train/test sets

```{python}
split = 0.8 # percentage of data to use for training
labels_map = labels_map.randomColumn('random') #set up a random column

training = labels_map.filter(ee.Filter.lt("random", split))
print('Training:')
print(training.getInfo())

testing = labels_map.filter(ee.Filter.gte("random", split))
print('Testing:')
print(testing.getInfo())
```

## Train the GTB model

```{python}
trainedGTB = (ee.Classifier.smileGradientTreeBoost(10).train(
  features = training,
  classProperty = 'byte_property',
  inputProperties = inputFeatures
))
```

## Evaluate the model

```{python}
confusionMatrixGTB = (testing
  .classify(trainedGTB)
  .errorMatrix(outputLabel, "classification"))
print('GTB Confusion Matrix:')
print(confusionMatrixGTB.getInfo())

acc_values_GTB = (confusionMatrixGTB
  .accuracy())
print("GTB Confusion Overall Accuracy: ", acc_values_GTB.getInfo())

```

# Apply model to image stack

## Helper functions

```{python}
#Calculate total area of AOI
def calc_area(feat):
  feat_area = feat.geometry().area()
  feat_area_ha = ee.Number(feat_area).divide(1e5)
  return feat.set('area_ha', feat_area_ha)

aoi_area = all_aois.map(calc_area)
print('total AOI area: ', aoi_area.getInfo())

# get CRS info
img_crs = ls_aoi.first().projection()
img_crsTrans = img_crs.getInfo().get('transform')

#function to apply the GTB model
def applyGTB(image):
  # Select the bands that correspond to the input features of the CART model
  imageFeatures = image.select(inputFeatures)
  missDate = image.get('missDate')
  # Classify the image using the trained GTB model
  classifiedImage = (imageFeatures
    .classify(trainedGTB)
    .set('missDate', missDate))
  return image.addBands(classifiedImage)


# save each value as its own band and mask
def extract_classes(image):
  cl = image.select('classification')
  cloud = cl.eq(0).rename('cloud').selfMask()
  openWater = cl.eq(1).rename('openWater').selfMask()
  lightNSSed = cl.eq(2).rename('lightNSSed').selfMask()
  OSSed = cl.eq(3).rename('OSSed').selfMask()
  dNSSed = cl.eq(4).rename('dNSSed').selfMask()
  classified = cl.gte(0).rename('classified').selfMask()
  img_addBand = (image.addBands(cloud)
    .addBands(openWater)
    .addBands(lightNSSed)
    .addBands(OSSed)
    .addBands(dNSSed)
    .addBands(classified))
  return img_addBand


def applyPerMissionDate(missDate):
  mission = ee.String(missDate).slice(0,9)
  date = ee.String(missDate).slice(10,20)

  short_stack = (ls_aoi
    .filter(ee.Filter.eq('SPACECRAFT_ID', mission))
    .filter(ee.Filter.eq('DATE_ACQUIRED', date)))

  oneMissDate = short_stack.mean()

  ls_miss_date_GTB = applyGTB(oneMissDate)

  ls_GTB_class = extract_classes(ls_miss_date_GTB)

  return (ls_GTB_class.set('missDate', missDate))


```

## consolidate stack by image date

```{python}
def addImageDate(image):
  mission = image.get('SPACECRAFT_ID')
  date = image.date().format('YYYY-MM-dd')
  missDate = ee.String(mission).cat('_').cat(ee.String(date))
  return image.set('missDate', missDate)

ls_aoi = ls_aoi.map(addImageDate)

# summarize by missionDate field
uniqueMissDate = ls_aoi.aggregate_array('missDate').distinct()

```


```{python}
def mosaicStack(missDate):
  md_GTB = applyPerMissionDate(missDate)
  return md_GTB

newStack_list = uniqueMissDate.map(mosaicStack)
newStack = ee.ImageCollection(newStack_list)

```

## Calculate area per image and export

```{python}
def calcArea(image):
  areaImage =  image.multiply(ee.Image.pixelArea()).divide(1e5)

  areaImage_light = areaImage.select([
    'classified',
    'cloud',
    'openWater',
    'lightNSSed',
    'OSSed',
    'dNSSed'
    ])

  area = areaImage_light.reduceRegions(
    collection = aoi_area,
    reducer = ee.Reducer.sum().forEachBand(areaImage),
    crs = img_crs,
    crsTransform = img_crsTrans
  )

  missDate = image.get('missDate')

  # Create a feature with the calculated area and properties
  a = area.first().set('missDate', missDate)

  return ee.FeatureCollection(a)


allAreas = newStack.map(calcArea).flatten()

for_export = allAreas
## export to drive
export_summary = (ee.batch.Export.table.toDrive(
  collection = allAreas,
  selectors = ['missDate', 'classified', 'cloud', 'openWater', 'lightNSSed','OSSed', 'dNSSed'],
  description = 'quick_gradientTreeBoost_landsat_stack_v2023-07-07',
  folder = 'eePlumB_classification',
  fileFormat = 'csv'))

export_summary.start()
```


## Export GeoTiffs to drive

Load task checker!

```{python}
##Function for limiting the max number of tasks sent to
#earth engine at one time to avoid time out errors
def maximum_no_of_tasks(MaxNActive, waitingPeriod):
  ##maintain a maximum number of active tasks
  ## initialize submitting jobs
  ts = list(ee.batch.Task.list())
  NActive = 0
  for task in ts:
     if ('RUNNING' in str(task) or 'READY' in str(task)):
         NActive += 1
  ## wait if the number of current active tasks reach the maximum number
  ## defined in MaxNActive
  while (NActive >= MaxNActive):
    time.sleep(waitingPeriod) # if reach or over maximum no. of active tasks, wait for 2min and check again
    ts = list(ee.batch.Task.list())
    NActive = 0
    for task in ts:
      if ('RUNNING' in str(task) or 'READY' in str(task)):
        NActive += 1
  return()

```

### Raster Export

Because the data storage types are different (int32 and float64), we have to
export the GTB output and the band data images separately. Note, each image
export takes 2-3 hours, resulting in a processing time of approximates 3 weeks.
Do not run these chunks until you're ready to export final files.

### GTB images

```{python}
date_length = len(uniqueMissDate.getInfo())

for d in range(date_length):
  md = uniqueMissDate.get(d)
  image = (newStack
    .filter(ee.Filter.eq('missDate', md))
    .first()
    .select('classification'))
  export_image = ee.batch.Export.image.toDrive(
    image = image,
    region = all_aois.geometry(),
    description = 'yugo_GTB_' + str(md.getInfo()),
    folder = 'yugo_GTB',
    scale = 30,
    crs = img_crs,
    maxPixels = 1e13)
  #Check how many existing tasks are running and take a break of 30 mins if it's >25
  maximum_no_of_tasks(10, 30*60)
  #Send next task.
  export_image.start()

```

### Band data images

```{python}
for d in range(date_length):
  md = uniqueMissDate.get(d)
  image = (newStack
    .filter(ee.Filter.eq('missDate', md))
    .first()
    .select(['SR_B2', 'SR_B3', 'SR_B4', 'SR_B5', 'SR_B6', 'SR_B7']))
  export_image = ee.batch.Export.image.toDrive(
    image = image,
    region = all_aois.geometry(),
    description = 'B2-B7_' + str(md.getInfo()),
    folder = 'superior_image_stack',
    scale = 30,
    crs = img_crs,
    maxPixels = 1e13)
  #Check how many existing tasks are running and take a break of 30 mins if it's >25
  maximum_no_of_tasks(25, 30*60)
  #Send next task.
  export_image.start()

```


```{r wrap, echo = F}
knitr::wrap_rmd('modeling/Yugo_Model.Rmd', width = 80, backup = NULL)
```
