---
title: "Landsat data availability"
author: "B Steele"
date: "2023-01-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('availability_checks/pySetup.R')
library(googledrive)

```

# Purpose

This script pulls MODIS availability from GEE and summarizes it for the designated AOI in western Lake Superior.

# Dependencies

[Install and initiate](https://cloud.google.com/sdk/docs/install) `gcloud`.

# Setup

Install python module.

```{python}
import ee
import datetime
import time
```

Authenticate `googledrive`

```{r}
drive_auth()
```

## Authenticate your Earth Engine Session

In the terminal in your Rstudio, authenticate your GEE login information. Do this by executing the following text in your zsh terminal.

`earthengine authenticate`

This will open a browser for you to login using your Google account associated with your EE login. If this fails and says that gcloud is not installed or recognized, make sure you're using a zsh terminal by clicking the dropdown next to 'Terminal 1', clicking properties, choosing 'New Terminal Opens With' zsh. For whatever reason (at least on a MacOS M1), a bash terminal will not execute the authentication process correctly. Additionally, there are commands (ee.Authenticate) in the Python GEE API, however it does not work in an .Rmd or .Qmd file code chunk.

Now we can initiate our EE session:

```{python}
ee.Initialize()

```

# Acquire data availability from GEE

Using GEE, summarize the proportion of area available for analysis within the AOI per day

## Load image collections and aoi

```{python}
#grab C2 SR/ST Landsat image collections and grab a consistent band and the qa pixel for this purpose.
l9 = ee.ImageCollection('LANDSAT/LC09/C02/T1_L2') \
  .select(['SR_B7', 'QA_PIXEL'])
l8 = ee.ImageCollection('LANDSAT/LC08/C02/T1_L2') \
  .select(['SR_B7', 'QA_PIXEL'])
l7 = ee.ImageCollection('LANDSAT/LE07/C02/T1_L2') \
  .select(['SR_B7', 'QA_PIXEL'])
l5 = ee.ImageCollection('LANDSAT/LT05/C02/T1_L2') \
  .select(['SR_B7', 'QA_PIXEL'])
l4 = ee.ImageCollection('LANDSAT/LT04/C02/T1_L2') \
  .select(['SR_B7', 'QA_PIXEL'])

ls = ee.ImageCollection(l4.merge(l5).merge(l7).merge(l8).merge(l8))

aoi = ee.FeatureCollection('projects/ee-ross-superior/assets/sup_pl_bl_small_aoi')

ls = ls.filterBounds(aoi)
```

## Cacluate AOI area

```{python}
aoi_area = aoi.geometry().area()
print(aoi_area.getInfo())
```

## Load Functions

#### clip: Function to clip image collection to AOI

```{python}
def clip(image):
  cl_im = image.clip(aoi)
  return cl_im

```

#### findWater_noClouds: Function to mask for water pixels and no clouds based on Landsat PIXEL_QA band.

Note, this is not the most robust way to perform either of these checks, but it will give us some idea of data availability

-   Bit 1: Dilated Cloud

-   Bit 3: Cloud

-   Bit 4: Cloud Shadow

-   Bit 5: Snow

-   Bit 7: Water

```{python}
def findWater_noClouds(image):
  qa = image.select('QA_PIXEL')
  water = qa.bitwiseAnd(1 << 7)
  dialatedclouds = qa.bitwiseAnd(1 << 1)
  clouds = qa.bitwiseAnd(1 << 3)
  shadow = qa.bitwiseAnd(1 << 4)
  snow = qa.bitwiseAnd(1 << 5)
  qaClouds = dialatedclouds.eq(1)\
    .where(clouds.eq(1), ee.Image(2)) \
    .where(shadow.eq(1), ee.Image(3)) \
    .where(snow.eq(1), ee.Image(4))
  maskClouds = qaClouds.eq(0)
  return image.updateMask(maskClouds).updateMask(water)

```

#### mosaicDate: Function to mosaic images from a given date

```{python}
def mosaicDate(date):
  oneDate = ls_aoi.filter(ee.String('DATE_ACQUIRED == ') \
    .cat(uniqueDates.getString(uniqueDates.indexOf(date))))
  additionalCollection = oneDate.mosaic()
  collectionOut = collectionOut.merge(additionalCollection)
  return collectionOut

```

#### addArea: Function to add area calculation per image

```{python}
def addArea(image):
  oneband = image.select('SR_B7') #grab any refl band 
  area = ee.Image.pixelArea() \
    .updateMask(oneband.mask()) \
    .reduceRegion(
      reducer = ee.Reducer.sum(), \
      geometry = aoi, \
      crs = img_crs, \
      crsTransform = img_crsTrans, \
      bestEffort = True
      ) 
  return image.set(area)

```

#### calcPro: Function to calculate the proportion of AOI available

```{python}
def calcPro(image):
  area = image.get('area')
  pro_area = ee.Number(area).divide(aoi_area)
  i = image.set('p_area', pro_area)
  return i

```

#### maximum_no_of_tasks: Define task counter

```{python}
# function to monitor jobs on GEE as to not send too many at once -- this is a function copied from the guts of AquaSat RS pulls, author is likely Simon Topp 

def maximum_no_of_tasks(MaxNActive, waitingPeriod):
  ##maintain a maximum number of active tasks
  time.sleep(10)
  ## initialize submitting jobs
  ts = list(ee.batch.Task.list())
  NActive = 0
  for task in ts:
     if ('RUNNING' in str(task) or 'READY' in str(task)):
         NActive += 1
  ## wait if the number of current active tasks reach the maximum number
  ## defined in MaxNActive
  while (NActive >= MaxNActive):
    time.sleep(waitingPeriod) # if reach or over maximum no. of active tasks, wait for 2min and check again
    ts = list(ee.batch.Task.list())
    NActive = 0
    for task in ts:
      if ('RUNNING' in str(task) or 'READY' in str(task)):
        NActive += 1
  return()

```

# Export LS Availability

## Declare timeframe and define CRS

Store crs and crsTransform for area calculation

```{python}
img_crs = ls.first().projection()
img_crsTrans = img_crs.getInfo()['transform']
```

## Pull all the functions together

```{python}
years = range(1983, 2023, 1)

for y in years:
  start = str(y) + '-04-01'
  end = str(y) + '-11-30'
  ls_filt = ls.filterDate(ee.Date(start), ee.Date(end))
  ls_qa = ls_filt.map(findWater_noClouds)
  ls_aoi = ls_qa.map(clip)
  
  allDates = ls_aoi.aggregate_array('DATE_ACQUIRED')
  uniqueDates = allDates.distinct().getInfo()
  
  collectionOut = ee.Image()
  
  for d in range(0, len(uniqueDates), 1):
    date = uniqueDates[d]
    datePlus1 = (datetime.datetime.strptime(date, '%Y-%m-%d') + datetime.timedelta(days = 1)).strftime('%Y-%m-%d')
    oneDate = ls_aoi.filterDate(ee.Date(date),ee.Date(datePlus1))
    w = oneDate.first().get('WRS_PATH')
    miss = oneDate.first().get('LANDSAT_PRODUCT_ID')
    if d > 0:
      collection = oneDate.mosaic().set({
        'wrs': w,
        'date': ee.String(date),
        'ls_miss': miss
      })
      addCollection = ee.ImageCollection(collection)
      collectionOut = collectionOut.merge(addCollection)
    else:
      collection = oneDate.mosaic().set({
        'wrs': w,
        'date': ee.String(date),
        'ls_miss': miss
      })
      collectionOut = ee.ImageCollection(collection)
      

  ls_area = ee.ImageCollection(collectionOut).map(addArea).map(calcPro)
  
  filename = 'superior_ls_availability_' + str(y) + '_v_' + datetime.datetime.today().strftime('%Y-%m-%d')
  
  file_out = ee.batch.Export.table.toDrive(
    collection = ls_area, 
    selectors = ['date', 'wrs', 'ls_miss', 'area', 'p_area'],
    description = filename,
    folder = 'superior_availability',
    fileFormat = 'csv')
  
  #Check how many existing tasks are running and take a break at a specified point 
  maximum_no_of_tasks(10, 120)
  file_out.start()

```

# Read in Files from Google Drive

```{r}
#get file list and ids
files = drive_ls('superior_availability/')
files = files[grepl('ls', files$name),]

#create a temp folder
dir.create('availability_checks/tmp')
tmpdir = 'availability_checks/tmp'
dr_down = function(fileid, filename) {
  drive_download(file = as_id(fileid), path = file.path(tmpdir, filename))
}

map2(files$id, files$name, dr_down)

```

## Collate Drive files together

```{r}
filelist = list.files(tmpdir)

read_file = function(file) {
  read.csv(file.path(tmpdir, file))
}

alldata = map_dfr(filelist, read_file)

alldata = alldata %>% rename(first_wrs = wrs)

datadir = 'data/inventory'
write.csv(alldata, file.path(datadir, 'superior_ls_inventory.csv'), row.names = F)
```

## Filter for 50% coverage

```{r}
filtered = alldata[alldata$p_area >= 0.5,]
# filtered = filtered[filtered$wrs == 26,] there are a few instances of multiple satellites imaging on a single day in neighboring paths, for now, let's just call this summary sufficient.

write.csv(filtered, file.path(datadir, 'superior_ls_inventory_gte50perc.csv'), row.names = T)
```
