---
title: "MODIS data availability"
author: "B Steele"
date: "2023-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('availability_checks/pySetup.R')
library(googledrive)
```

# Purpose

This script pulls MODIS availability from GEE and summarizes it for the designated AOI in western Lake Superior.

# Dependencies

[Install and initiate](https://cloud.google.com/sdk/docs/install) `gcloud`.

# Setup

Install python module.

```{python}
import ee
from datetime import datetime
import time
```

Authenticate `googledrive`

```{r}
drive_auth()
```

## Authenticate your Earth Engine Session

In the terminal in your Rstudio, authenticate your GEE login information. Do this by executing the following text in your zsh terminal.

`earthengine authenticate`

This will open a browser for you to login using your Google account associated with your EE login. If this fails and says that gcloud is not installed or recognized, make sure you're using a zsh terminal by clicking the dropdown next to 'Terminal 1', clicking properties, choosing 'New Terminal Opens With' zsh. For whatever reason (at least on a MacOS M1), a bash terminal will not execute the authentication process correctly. Additionally, there are commands (ee.Authenticate) in the Python GEE API, however it does not work in an .Rmd or .Qmd file code chunk.

Now we can initiate our EE session:

```{python}
ee.Initialize()
```

# Acquire data availability from GEE

## Load image collections and aoi

```{python}
aqua = ee.ImageCollection('MODIS/061/MYD09GA')
terra = ee.ImageCollection("MODIS/061/MOD09GA")

aoi = ee.FeatureCollection('projects/ee-ross-superior/assets/sup_pl_bl_small_aoi')
```

## Cacluate AOI area

```{python}
aoi_area = aoi.geometry().area()
print(aoi_area.getInfo())
```

## Load Functions

#### Helper function to extract the values from specific bits.

The input parameter can be a ee.Number() or ee.Image(). Code adapted from <https://gis.stackexchange.com/a/349401/5160>

```{python}
def bitwiseExtract(input, fromBit, toBit): 
  maskSize = ee.Number(1).add(toBit).subtract(fromBit)
  mask = ee.Number(1).leftShift(maskSize).subtract(1)
  return input.rightShift(fromBit).bitwiseAnd(mask)

```

#### Function to clip image collection to AOI

```{python}
def clip(image):
  cl_im = image.clip(aoi)
  return cl_im

```

#### Function to mask for water pixels based on MODIS QA state band.

**Bits 3-5: Land/water flag**

0: Shallow ocean

1: Land

2: Ocean coastlines and lake shorelines

3: Shallow inland water

4: Ephemeral water

5: Deep inland water

6: Continental/moderate ocean

7: Deep ocean

```{python}
def find_water(image):
  state = image.select('state_1km')
  shallow_water = bitwiseExtract(state, 3, 5).eq(3)
  deep_water = bitwiseExtract(state, 3, 5).eq(5)
  land = bitwiseExtract(state, 3, 5).eq(1)
  water_mask = shallow_water.eq(1)\
    .Or(deep_water.eq(1))\
    .And(land.neq(1))
  return image.updateMask(water_mask)

```

#### Function to mask clouds

**Bits 0-1: Cloud state**

0: Clear

1: Cloudy

2: Mixed

3: Not set, assumed clear

**Bit 2: Cloud shadow**

0: No

1: Yes

```{python}
def maskCloud(image):
  state = image.select('state_1km')
  clear = bitwiseExtract(state, 0, 1).eq(0).Or(bitwiseExtract(state, 0,1).eq(3))
  no_shad = state.bitwiseAnd(1 << 2).neq(1)
  cloud_mask = clear.eq(1).And(no_shad.eq(1))
  return image.updateMask(cloud_mask)

```

#### Function to add area calculation per image

```{python}
def addArea(image):
  oneband = image.select('sur_refl_b01') #grab any refl band 
  area = ee.Image.pixelArea() \
    .updateMask(oneband.mask()) \
    .reduceRegion(
      reducer = ee.Reducer.sum(), \
      geometry = aoi, \
      crs = img_crs, \
      crsTransform = img_crsTrans, \
      bestEffort = True
      ) 
  return image.set(area)

```

#### Function to calculate the proportion of AOI available

```{python}
def calcPro(image):
  area = image.get('area')
  pro_area = ee.Number(area).divide(aoi_area)
  i = image.set('p_area', pro_area)
  return(i)

```

## Declare timeframe and define CRS

Store crs and crsTransform for area calculation

```{python}
img_crs = terra.first().select('sur_refl_b01').projection()
img_crsTrans = img_crs.getInfo()['transform']

```

Define the function to walk through the years of images

```{python}

def defineAquaCollection(year):
  start_date = str(year) + '-04-01'
  end_date = str(year) + '-11-01'
  imgcoll = aqua \
    .filter(ee.Filter.date(start_date, end_date))
  return imgcoll


def defineTerraCollection(year):
  start_date = str(year) + '-04-01'
  end_date = str(year) + '-11-01'
  imgcoll = terra \
    .filter(ee.Filter.date(start_date, end_date))
  return imgcoll
  
  
def calcProArea(image):
  #extract to AOI
  coll_aoi = clip(image)
  #find water
  coll_water = find_water(coll_aoi)
  #mask clouds
  coll_qual = maskCloud(coll_water)
  #calculate area and proportion of area
  coll_area = calcPro(addArea(coll_qual))
  return coll_area

```

Define task counter

```{python}
# function to monitor jobs on GEE as to not send too many at once -- this is a function copied from the guts of AquaSat RS pulls, author is likely Simon Topp 

def maximum_no_of_tasks(MaxNActive, waitingPeriod):
  ##maintain a maximum number of active tasks
  time.sleep(10)
  ## initialize submitting jobs
  ts = list(ee.batch.Task.list())
  NActive = 0
  for task in ts:
     if ('RUNNING' in str(task) or 'READY' in str(task)):
         NActive += 1
  ## wait if the number of current active tasks reach the maximum number
  ## defined in MaxNActive
  while (NActive >= MaxNActive):
    time.sleep(waitingPeriod) # if reach or over maximum no. of active tasks, wait for 2min and check again
    ts = list(ee.batch.Task.list())
    NActive = 0
    for task in ts:
      if ('RUNNING' in str(task) or 'READY' in str(task)):
        NActive += 1
  return()

```

# Export Terra Availability

```{python}
years = range(2000, 2023, 1)

for year in years:
  terra_collection_sub = defineTerraCollection(year)
  terra_calc_sub = terra_collection_sub.map(calcProArea)
  filename = 'superior_terra_availability_' + str(year) + '_v' + datetime.today().strftime('%Y-%m-%d')
  file_out = ee.batch.Export.table.toDrive(
    collection = terra_calc_sub, 
    selectors = ['system:index', 'area', 'p_area'],
    description = filename,
    folder = 'superior_availability',
    fileFormat = 'csv')
  #check number of jobs
  maximum_no_of_tasks(10,120) #if there are 10 tasks pending, wait 2 minutes, this is very conservative, you could easily increase tasks or decrease waiting time to slightly quicken this process
  file_out.start()

```

# Export Aqua Availability

```{python}
years = range(2002, 2023, 1)

for year in years:
  aqua_collection_sub = defineAquaCollection(year)
  aqua_calc_sub = aqua_collection_sub.map(calcProArea)
  filename = 'superior_aqua_availability_' + str(year) + '_v' + datetime.today().strftime('%Y-%m-%d')
  file_out = ee.batch.Export.table.toDrive(
    collection = aqua_calc_sub, 
    selectors = ['system:index', 'area', 'p_area'],
    description = filename,
    folder = 'superior_availability',
    fileFormat = 'csv')
  #check number of jobs
  maximum_no_of_tasks(10,120) #if there are 10 tasks pending, wait 2 minutes, this is very conservative, you could easily increase tasks or decrease waiting time to slightly quicken this process
  file_out.start()

```

# Read in Files from Google Drive

```{r}
#get file list and ids
files = drive_ls('superior_availability/')
files = files[grepl('aqua', files$name) | grepl('terra', files$name) ,]

#create a temp folder
dir.create('availability_checks/tmp')
tmpdir = 'availability_checks/tmp'
dr_down = function(fileid, filename) {
  drive_download(file = as_id(fileid), path = file.path(tmpdir, filename))
}

map2(files$id, files$name, dr_down)

```

## Collate Drive files together

```{r}
filelist = list.files(tmpdir)

read_file = function(file) {
  df = read.csv(file.path(tmpdir, file)) %>% 
    mutate(sat = unlist(str_split(file, '_'))[2],
    year = unlist(str_split(file, '_'))[4])
  df
}

alldata = map_dfr(filelist, read_file)

datadir = 'data/inventory'
write.csv(alldata, file.path(datadir, 'superior_modis_inventory.csv'), row.names = F)
```

## Filter for 50% coverage

```{r}
filtered = alldata[alldata$p_area >= 0.5,]

write.csv(filtered, file.path(datadir, 'superior_modis_inventory_gte50perc.csv'), row.names = T)
```
